\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{breqn}

\newcommand{\la}{\left\langle}
\newcommand{\ra}{\right\rangle}
\newcommand{\beq}{\begin{eqnarray}}
\newcommand{\eeq}{\end{eqnarray}}

\newcommand{\append}{\leftarrow{\mkern-18mu}{+}\:}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

\title{DRAFT: A Family of Sparse, Vertex-Transitive Graphs}
\author{Edward L. Platt}
\date{August 2015}

\begin{document}

\maketitle

\section{Motivation}
In networks modeling interactions between agents (e.g. people, computers),
the agents often have a limited capacity to form interactions.
If the vertex degrees grow with the size of the network,
the number of interactions will exceed the capacity of the agents,
limiting the scale of the network.
Sparse networks, such as regular trees, are not subject to this limitation on scaling.
However, trees impose a hierarchical relationship between agents.
In some applications, an egalitarian relationship between agents is desirable,
formally described as vertex transitivity.
In this paper we construct an infinite family $G_n$
of sparse ($|E| = O(|V|\log(\log |V|))$,
small-diameter ($d = O(\log |V|)$), vertex-transitive graphs.

\section{Construction}

We recursively construct a family of vertex-transitive graphs
$G_n = \la V_n, E_n \ra$.
The vertices $v \in V_n$ are labeled by an $n$-sequence of integers.
We define $G_0$ and $G_1$ as:
\beq
V_0 &=& \{ \la\ra \}, \\
E_0 &=& \{ \}, \\
V_1 &=& \{\la{0}\ra,\la{1}\ra\}, \\
E_1 &=& \left\{\{\la{0}\ra,\la{1}\ra\}\right\}.
\eeq
We construct subsequent vertex sets from copies of the previous set,
with each copy having a different integer appended to its vertex labels:
\beq
\label{eq:vn}
V_{n+1} &=& \bigcup_{k = 0}^{C_n}
\left\{ v | w \in V_n \land v = w \append k \right\},
\\
C_n &\equiv& |V_n|,
\eeq
where $w \append k$ denotes appending element $k$ to the end of sequence $w$.
We note that a one-to-one mapping $z_n$ exists between the vertices of $G_n$ and the
integers from $0$ to $C_n - 1$:
\beq
z_n(v) &=& \sum_{i=0}^{n-1} C_i v_i.
\eeq
We define even and odd subsets of $V_n$:
\beq
A_n &=& \{ v | v \in V_n \land v_0 = 1 \}, \\
B_n &=& \{ v | v \in V_n \land v_0 = 0 \}.
\eeq
We also define the following mappings on the vertices of $G_n$:
\beq
(\phi_k(v))_i
&=&
\begin{cases}
v_i + 1 \mbox{ mod } (C_i + 1) & \mbox{if } i = k,
\\
v_i & \mbox{otherwise},
\end{cases}
\\
(\psi_n(v))_i
&=&
\begin{cases}
v_i + 1 \mbox{ mod } (C_i + 1)
& \mbox{if } i = 0 \lor \forall j < i: v_j = C_j, \\
v_i & \mbox{otherwise},
\end{cases}
\\
\theta_n(v)
&=&
\begin{cases}
\psi_n(v) & \mbox{if } v \in A_n, \\
\psi_n^{-1}(v) & \mbox{if } v \in B_n,
\end{cases}
\eeq
with $i, k < n$.

To construct the edges of $G_{n+1}$ we define a ``shortcut'' function $s_n(j, k)$
with $j \in \{0, 1, C_n - 1\}$ and $k \in \{0, 1, 2, \ldots, C_n\}$ which
determines the interconnections between copies of $G_n$.
\beq
\label{eq:shortcut}
s_{n}(j, k)
&=&
\begin{cases}
k + 1 \ \mbox{mod } (C_n + 1)
&
\mbox{if } j = C_n - 1,
\\
k - 1 \ \mbox{mod } (C_n + 1)
&
\mbox{if } j = 0,
\\
k + j + 1 \ \mbox{mod } (C_n + 1)
&
\mbox{if } j \in \{1, 3, \ldots, C_n - 3\},
\\
k - j \ \mbox{mod } (C_n + 1)
&
\mbox{if } j \in \{2, 4, \ldots, C_n - 2\}.
\end{cases}
\eeq
We also define a parity selector function on the edges $e \in E_n$:
\beq
p_x(e)
&=&
v \quad \mbox{s.t. } v \in e \land v_0 = x.
\eeq
The edges of $G_{n+1}$ are then given by:
\beq
\label{eq:edgesrecursive}
R_{n+1}
&=&
\bigcup_{k=0}^{C_n}
\bigcup_{e \in E_n} \{ p_0(e) \append k, p_1(e) \append k\},
\\
\label{eq:edgesshortcut}
S_{n+1}
&=&
\bigcup_{k=0}^{C_n}
\bigcup_{v \in A_n} \{ v \append k, \theta_n(v) \append s_n(z_n(v), k) \},
\\
&=&
\bigcup_{k=0}^{C_n}
\bigcup_{v \in B_n} \{ v \append k, \theta_n(v) \append s_n(z_n(v), k) \},
\\
E_{n+1}
&=&
R_{n+1} \cup S_{n+1},
\eeq
noting that $S_{n+1}$ can be written in terms of either the odd vertices $A_n$
or the even vertices $B_n$.
Eq. (\ref{eq:edgesrecursive}) replicates the edges of
$G_n$ among subsets of the vertices of $G_{n+1}$,
while Eq. (\ref{eq:edgesshortcut})
creates one edge between each pair of the subsets.

\section{Properties of $G_n$}

\begin{lemma}
\label{lem:regular}
The graph $G_n$ is $n$-regular for all $n \geq 0$.
\end{lemma}
\begin{proof}
We proceed using induction on $n$.
The base case $G_1$ is $1$-regular by inspection.
In the inductive case $G_{n+1}$,
Eq. (\ref{eq:edgesrecursive})
reproduces the edges of $G_n$,
which is $n$-regular by induction,
contributing $n$ to the degree of each vertex.
Eq. (\ref{eq:edgesshortcut}) adds one to the degree of each odd vertex.
As $\theta_n$ is a bijective map between odd and even vertices,
Eq. (\ref{eq:edgesshortcut}) also adds one to the degree of each even vertex,
giving a total degree of $n + 1$.
\end{proof}

\begin{lemma}
The number of vertices and edges of the graph $G_n$ are given by the recurrence
relations:
\beq
C_0 = |V_0| &=& 1, \\
\label{eq:cnrec}C_{n} = |V_{n}| &=& C_{n-1} (C_{n-1} + 1), \\
|E_n| &=& \frac{n}{2} C_n.
\eeq
\end{lemma}
\begin{proof}
The vertex set of $G_n$, given by Eq. (\ref{eq:vn}), is a union of
$C_{n-1} + 1$ sets.
The vertex labels within each set all end with the same element,
and this element is unique to each set.
The sets are thus disjoint.
Each set contains $C_{n-1}$ elements, giving $C_{n-1}(C_{n-1} + 1)$
elements.
By Lemma \ref{lem:regular}, $G_n$ is $n$-regular and has
$\frac{n}{2}|V_n| = \frac{n}{2}C_n$ edges.
\end{proof}
\begin{lemma}
\label{lem:phiauto}
The mapping $\phi_{n-1}$ is an automorphism of $G_n$.
\end{lemma}
\begin{proof}
Let $v$ be a vertex in $V_n$.
$\phi_{n-1}$ has no effect on $v_0^{n-2}$ so it preserves the edges $R_n$,
defined in Eq. (\ref{eq:edgesrecursive}).
The vertex $v$ has exactly one other edge, from $S_n$ defined in
Eq. (\ref{eq:edgesshortcut}):
\beq
e &=& \{ v, \theta_n(v_0^{n-2} \append s_n(z_n(v_0^{n-2}), v_{n-1}) \}.
\eeq
Applying $\phi_{n-1}$ to both vertices gives:
\beq
\tilde{e} &=&
\{ v_0^{n-2} \append (v_{n-1} + 1) \mbox{ mod } (C_{n-2} + 1), \\
&& \quad \theta_n(v_0^{n-2} \append
s_n(z_n(v_0^{n-2}), (v_{n-1} + 1) \mbox{ mod } (C_{n-2} + 1) ) \},
\\
&=&
\{ v_0^{n-2} \append \tilde{k},
\theta_n(v_0^{n-2} \append s_n(z_n(v_0^{n-2}), \tilde{k}) \},
\eeq
where $\tilde{k} = (v_{n-1} + 1) \mbox{ mod } (C_{n-2} + 1)$.
The edge $\tilde{e}$ is also a member of $S_n$,
showing that $\phi_{n-1}$ permutes the elements of $S_n$,
and preserves all edges in $E_n$.
\end{proof}
\begin{theorem}
The graphs $G_n$ are vertex-transitive for all $n \geq 0$.
\end{theorem}
\begin{proof}
The graph $G_0$ is trivially vertex-transitive.
For higher order graphs,
we recursively construct a mapping $\Phi_n$
that maps an arbitrary vertex $v \in V_n$ to
an arbitrary vertex $\tilde{v} \in V_n$ and preserves the edges in $E_n$.
Let $\delta_i = \tilde{v}_i - v_i \mbox{ mod } (C_i + 1)$.
For the subsequence $v_0^0$, the map $\Phi_1 = \phi_0^{\delta_0}$
achieves the desired mapping and preserves the edges in $E_1$ by
Lemma \ref{lem:phiauto}.
Given an edge-preserving mapping $\Phi_n$
from $v_0^{n-1}$ to $\tilde{v}_0^{n-1}$,
we can construct $\tau_{n}$.
We define mappings $\pi_n$ and $\tau_{n}$:
\beq
\pi_n(0) &=& 0
\\
\pi_n(s_n(v_0^{n-1}, 0))
&=& s_n(\Phi_n(v_0^{n-1}), 0)
\\
\tau_{n}(v)
&=& \Phi_{n}(v_0^{n-1}) \append \pi_n(v_n).
\eeq
For the recursive edges defined in Eq. (\ref{eq:edgesrecursive})
$\{v_0^n, w_0^n\} \in R_{n+1}$:
\beq
\{ \tau_{n}(v_0^n), \tau_{n}(v_0^n)
&=&
\{ \Phi_n(v_0^n) \append \pi_n(k), \Phi_n(w_0^n) \append \pi_n(k)\}
\\ &=&
\{ \tilde{v}_0^n \append \tilde{k}, \Phi_n(\tilde{w}_0^n) \append \tilde{k}\},
\eeq
which is simply another member of $R_{n+1}$.
As $\tau_{n}$ is bijective, it permutes the edges of $R_{n+1}$.
For the shortcut edges defined in Eq. (\ref{eq:edgesshortcut})
$\{v_0^n, w_0^n\} \in S_{n+1}$, when $v_n=0$:
\beq
\{\tau_{n}(v_0^n), \tau_{n}(w_0^n)\}
&=&
\{ \Phi_n(v_0^n) \append \pi_n(0),
\\ &&
\quad \Phi_n(w_0^n)) \append \pi_n(s_n(z_n(v_0^n), 0)) \}
\\ &=&
\{ \tilde{v}_0^n \append 0,
\tilde{w}_0^n \append s_n(z_n(\Phi_n(v_0^n)), 0) \}
\\ &=&
\{ \tilde{v}_0^n \append 0,
\theta_n(\tilde{v}_0^n) \append s_n(z_n(\tilde{v_0^n}), 0) \},
\eeq
showing that $\tau_{n}$ permutes the subset of edges in $S_{n+1}$ for which
$v_n=0$.
Consequently, because $\phi_n$ is an automorphism of $G_{n+1}$
by Lemma \ref{lem:phiauto}, $\tau_{n}$ also permutes the edges of the
subsets for which $(\phi_n^{-k}(v_0^n))_n=0$.
Those edges are the subset of $S_{n+1}$ for which $v_n=k$,
showing that $\Phi_n{n+1}$ permutes all edges within $S_{n+1}$ and
$E_{n+1}$.
We finish by applying the automorphism $\phi_n^{\delta_n}$:
\beq
\Phi_{n+1}(v) &=& \phi_n^{\delta_n}(\tau_{n}(v)),
\eeq
completing the mapping from $\tilde{v}$ to $\tilde{w}$.
\end{proof}
\section{Acknowledgements}
% Joe Nievelt
% Aram Harrow
% Tony Garnock-Jones
% Cesar Hidalgo
% Grant Schoenebeck
\section{To Do}
\begin{enumerate}
\item{Change Eq. (\ref{eq:shortcut}) to $C_n - j + k \ (\mbox{mod } C_n + 1)$ to reduce cases.}
\item{Prove degree as function of $n$ following \cite{aho1973}.}
\end{enumerate}
\bibliographystyle{plain}
\bibliography{vtgraph}

\end{document}
